---
title: "8_Association_Rule_Mining"
output: pdf_document
date: "2023-08-13"
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
groceries = read.csv('~/STA380-Exercise/datafile/groceries.txt', header = FALSE)
```

```{r}
library(arules)

# Replace empty cells with NA
groceries[groceries == ""] <- NA

# Create baskets by removing NA values
baskets <- lapply(1:nrow(groceries), function(row) {
  basket <- unlist(groceries[row, ])
  basket <- basket[!is.na(basket)]
  basket
})

# Convert baskets to transactions
gtrans <- as(baskets, "transactions")

# Display summary of transactions
summary(gtrans)
```

The data consists of 15,296 shopping baskets (transactions) and 169
unique items (columns). The density of the data indicates that, on
average, each transaction contains approximately 1.68% of the total
available items.

```{r}

# Item frequency plot
itemFrequencyPlot(
  gtrans, topN = 20,
  type = 'relative',
)
```

Frequent Items: Among the most frequent items, we find that "whole
milk," "other vegetables," "rolls/buns," "soda," and "yogurt" dominate
the baskets. These items are commonly purchased by the customers.

Transaction Length Distribution: The distribution of transaction lengths
provides an overview of the number of items per basket. It appears that:

3,485 baskets contain 1 item, 2,630 baskets contain 2 items 2,102
baskets contain 3 items, 7,079 baskets contain 4 items Item

Labels: We also have additional information about some of the items. For
instance, we can see examples of items like "abrasive cleaner,"
"artificial sweetener," and "baby cosmetics."

### Finding interesting association rules

```{r}
# Perform association rule mining using Apriori algorithm
grocr <- apriori(gtrans, 
                 parameter = list(support = 0.005, confidence = 0.1, maxlen = 4))

# Display summary of the mined rules
summary(grocr)
```

Rule Characteristics: The Apriori algorithm discovered a total of 118
association rules, characterized by varying lengths of items on both the
left-hand side (lhs) and right-hand side (rhs) of the rules. These rules
provide insights into how items are related in the context of grocery
shopping baskets.

Support: The minimum support for an item set is approximately 0.5%,
indicating that a rule is relevant if it appears in at least this
proportion of transactions.

Confidence: The minimum confidence threshold is set at 10%, meaning that
a rule should have a confidence level of at least this value to be
considered significant.

Coverage: The coverage reflects the proportion of transactions covered
by the rule.

Lift: The lift measures the strength of the association between the
items in the rule. A lift greater than 1.0 indicates a positive
association.

```{r}
# Filter rules based on lift threshold
high_lift_rules <- subset(grocr, subset = lift > 2.5)

# Display and interpret the high lift rules
inspect(high_lift_rules)

```

These high lift association rules provide a glimpse into intriguing
purchasing patterns. Connection between onions and root vegetables
suggests that customers who buy onions are likely to purchase other root
vegetables and other vegetables as well. Beef buyers also show an
interest in citrus fruit and root vegetables.

Picking Thresholds: I set a minimum confidence threshold of 10% and a
minimum lift threshold of 2.5. Confidence Threshold: The confidence
threshold of 10% was chosen to identify rules where the consequent (rhs)
appears in at least 10% of transactions containing the antecedent (lhs).
Lift Threshold: A lift value of 2.5 implies that the items are 2.5 times
more likely to be bought together than expected by chance.

```{r}
# Load necessary packages for visualization
library(arulesViz)

# Visualize the high lift rules, including weaker associations
plot(high_lift_rules, method = "graph", control = list(type = "items", shading = "lift"))

```

In the graph, each item is represented as a node (circle). These nodes
are interconnected by edges (lines), illustrating the linkages between
items based on the high lift association rules.

The direction of the arrows on the edges indicates the direction of
association between items. For example, customers who buy whole milk are
more likely to also purchase vegetables.
